Notes

### 2022-09-22 notes/notes.tex
append ideas to further work

To correct for non-detections, we apply the transformation $h$ from \cite{deMeutter2022uncertainty}. This is given by the continuously differentiable function
$$
h(y) = \begin{cases}
    \frac{y^2}{4MDC} + MDC &\quad \text{if } y \le 2MDC \\
    y &\quad \text{else}
\end{cases}
.$$
This transformation maps the interval $[0, MDC]$ onto $[MDC, 2MDC]$. For observations that are significantly larger than $MDC$, this is approximately the identity. If $y$ and the simulated value $\hat{y}$ are both larger than $2MDC$, then $h(\hat{y}) / h(y) = \hat{y} / y$. If $y = 0$ because the concentration is not detectable, then we want a simulated value $\hat{y}$ such that $h(\hat{y}) \approx MDC$. In both cases, we can use the value of $h(\hat{y}) / h(y)$ as an indication of the accuracy of the simulation.

@@

\section{Further work}
\begin{itemize}
    \item For certain emission sources in certain periods, there is an expected correlation between the emission at time $t$ and at time $t + 1, t + 2, \dots$. We can add a constraint on the autocorrelation.
    \item Time-correlation could be measured by parameterising the model as follows: 
    \begin{align*}
        x_1^{2 \text{ Jan } 2014} &= \alpha_1^{(1)} x_1^{1 \text{ Jan } 2014} \\
        x_1^{3 \text{ Jan} 2014} &= \alpha_2^{(1)} x_1^{2 \text{ Jan } 2014} \\
        &\vdots
    \end{align*}
    with a constraint like $\alpha_{\min} \le \prod_{t=1}^T \alpha_t^{(j)} \le \alpha_{\max}$ for all $T$ and all $j = 1,\dots,m$. However, this would be difficult to implement. In addition, the complexity of inequality constrained optimisation can be combinatorial in the number of constraints if an active set method is used \cite{Nocedal2006}.

@@

    \item Standard software libraries support optimisation with constraints. This is an alternative to regularisation and may be more interpretable.
    Suppose that the number of components in $\tilde{x}$ is $m T$ (where $m$ is the number of sources and $T$ is the number of days).
    If we want to bound the average of $(\log w_{ij})^2$ by some value $\delta^2$, then we have the constrained optimisation problem
    \begin{align*}
        \min\quad & \frac{1}{2} \norm{\log h.(\hat{y})}^2 \\
        \text{s.t.}\quad & \frac{1}{mT} \norm{\log W}^2 \le \delta^2
    \end{align*} 
    which can be solved using similar ideas as those developed above. The prototypical constrained optimisation method (such as an active set method) would iterate towards a critical point of the Lagrangian
    $$
    \mathcal{L}(W, \lambda) = \frac{1}{2} \norm{\log h.(\hat{y})}^2 + \lambda\left(\frac{1}{mT} \norm{\log W}^2 - \delta^2\right)
    $$
    in which case the computational cost would be comparable to optimising \cref{eq: cost function regularised}. A more detailed explanation of constrained optimisation and the Lagrangian is found in \cite{Nocedal2006}.
\end{itemize}

>>>>

Sometimes, the concentration at a measurement station is below the minimum detectable concentration (MDC). The corresponding measurements are $y_i = 0$.
To account for non-detections, we apply the transformation $h$ from \cite{deMeutter2022uncertainty}. This is given by the continuously differentiable function
$$
h(y) = \begin{cases}
    \frac{y^2}{4MDC} + MDC &\quad \text{if } y \le 2MDC \\
    y &\quad \text{else}
\end{cases}
.$$
This transformation maps the interval $[0, MDC]$ onto $[MDC, 2MDC]$. For observations that are significantly larger than $MDC$, this is approximately the identity. If $y$ and the simulated value $\hat{y}$ are both larger than $2MDC$, then $h(\hat{y}) / h(y) = \hat{y} / y$. If $y = 0$ because the concentration is not detectable, then we want a simulated value $\hat{y}$ such that $h(\hat{y}) \approx MDC$. In both cases, we can use the value of $h(\hat{y}) / h(y)$ as an indication of the accuracy of the simulation.

@@

\section{Further work}
\begin{itemize}
    \item Standard software libraries support optimisation with constraints. This is an alternative to regularisation and may be more interpretable.
    Suppose that the number of components in $\tilde{x}$ is $m T$ (where $m$ is the number of sources and $T$ is the number of days).
    If we want to bound the average of $(\log w_{ij})^2$ by some value $\delta^2$, then we have the constrained optimisation problem
    \begin{align*}
        \min\quad & \frac{1}{2} \norm{\log h.(\hat{y})}^2 \\
        \text{s.t.}\quad & \frac{1}{mT} \norm{\log W}^2 \le \delta^2
    \end{align*} 
    which can be solved using similar ideas as those developed above. The prototypical constrained optimisation method (such as an active set method) would iterate towards a critical point of the Lagrangian
    $$
    \mathcal{L}(W, \lambda) = \frac{1}{2} \norm{\log h.(\hat{y})}^2 + \lambda\left(\frac{1}{mT} \norm{\log W}^2 - \delta^2\right)
    $$
    in which case the computational cost would be comparable to optimising \cref{eq: cost function regularised}. A more detailed explanation of constrained optimisation and the Lagrangian is found in \cite{Nocedal2006}.
    \item Currently, there is no constraint saying that, for every emission source $j$, the geometric mean of the emission should be the reported average $x_j$. This can be enforced with a constraint like $\prod_{t} w_{jt} = 1$ for all sources $j=1,\dots,m$. With the parametrisation $w = e^v$, this is equivalent to the linear constraint $\sum_{t} v_{jt} = 0$. Since this is a linear constraint, it can be handled relatively easily.
    \item For certain emission sources in certain periods, there is an expected correlation between the emission at time $t$ and at time $t + 1, t + 2, \dots$. We can add a constraint on the autocorrelation.
    \item Time-correlation could be measured by parameterising the model as follows: 
    \begin{align*}
        x_1^{2 \text{ Jan } 2014} &= \alpha_1^{(1)} x_1^{1 \text{ Jan } 2014} \\
        x_1^{3 \text{ Jan} 2014} &= \alpha_2^{(1)} x_1^{2 \text{ Jan } 2014} \\
        &\vdots
    \end{align*}
    with a constraint like $\alpha_{\min} \le \prod_{t=1}^T \alpha_t^{(j)} \le \alpha_{\max}$ for all $T$ and all $j = 1,\dots,m$. However, this would be difficult to implement. In addition, the complexity of inequality constrained optimisation can be combinatorial in the number of constraints if an active set method is used \cite{Nocedal2006}.

@@

\end{itemize}
